---
title: "Batch effects"
author: "Şebnem Perihan Gök"
format: html
editor: visual
---

# Introduction

The purpose of this tutorial is to give a detailed information on identifying batch effects and how to remove them.

# Methods

## Data

To download the raw affymetryx data, [use this link](https://ndownloader.figshare.com/files/4862323) and save it.

First create a `data` folder:

``` bash
mkdir data
```

Save the affymetrix CEL files in the data

## Libraries

We will use these libraries:

```{r}
library(affy)
library(genefilter)
```

## Preprocess data

Load the affmetryx files:

```{r}
tab=read.csv("data/bladdertab.csv", as.is=TRUE)
eset=justRMA(filenames=tab$filename, celfile.path="data/")
```
## Obtaining the gene expression matrix

remember you downloaded library(genefilter), you are using it here, 
```{r}
outcome=tab[,8]
bt=tab[,5]
Index=which(outcome=="sTCC")
Cplus=grep("CIS",bt[Index])
outcome[Index]="sTCC-CIS"
outcome[Index[Cplus]]="sTCC+CIS"
outcome[49:57]<-"Biopsy"
mat=exprs(eset)

dates=vector("character",ncol(mat))
for(i in seq(along=dates)){
    tmp=affyio::read.celfile.header(file.path("bladdercels",tab$filenam[i]),info="full")$DatHeader
  dates[i]=strsplit(tmp,"\ +")[[1]][8]
}
dates=as.Date(dates,"%m/%d/%Y")
batch=dates-min(dates);
batch=as.numeric(cut(batch,c(-1,10,75,200,300,500))) 
```

## obtaining the R object

```{r} 
save(mat,outcome,batch,dates,file="bladder-cancer.rda")
```










